Convolutional Neural Networks and Volcano Plots: Screening and Prediction of Two-Dimensional Single-Atom Catalysts for CO\textsubscript{2} Reduction Reactions

### 2024-02-06 paper/arxiv-submission/main.tex
merge supplement sections
remove TODOs
add comment back anyway
add header comment
resolve: Warning: Token not allowed in a PDF string (Unicode): (hyperref)

\documentclass[a4paper, 12pt, titlepage]{article}

% TODO: remove unused?
\usepackage{amsmath}
\usepackage{array}
\usepackage{bookmark}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{cleveref}
\usepackage{datetime2}
\usepackage{enumitem}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{nicematrix}
\usepackage{tablefootnote}
\usepackage{textcomp}
\usepackage{siunitx}
\usepackage{xr}
\usepackage[affil-it]{authblk}
\usepackage[english]{babel}
\usepackage[version=4]{mhchem}

@@

\externaldocument{supp}  % TODO: remove this?

@@

\newpage
    \title{Convolutional Neural Networks and Volcano Plots:
    Screening and Prediction of Two-Dimensional Single-Atom Catalysts
    for CO\textsubscript{2} Reduction Reactions}

@@

\newpage
\begin{abstract}
    Single-atom catalysts (SACs) have emerged as pivotal frontiers for catalyzing a myriad of chemical reactions,
    yet the diverse combinations of active elements and support materials, the nature of coordination environments,
    elude traditional methodologies in searching optimal SAC systems with superior catalytic performance.
    Herein, by integrating multi-branch Convolutional Neural Network (CNN) analysis models
    to hybrid descriptor based activity volcano plot,
    two-dimensional (2D) SAC system composed of diverse metallic single atoms anchored on six type of 2D supports,
    including graphitic carbon nitride (g-C\textsubscript{3}N\textsubscript{4}), nitrogen-doped graphene,
    graphene with dual-vacancy, black phosphorous, boron nitride (BN), and C\textsubscript{2}N,
    are screened for efficient CO\textsubscript{2}RR.
    Starting from establishing a correlation map between the adsorption energies of
    intermediates and diverse electronic and elementary descriptors,
    sole singular descriptor lost magic to predict catalytic activity, including d-band center.
    Deep learning method utilizing multi-branch CNN model therefore was employed,
    using 2D electronic density of states (eDOS) as input to predict adsorption energies.
    Hybrid-descriptor enveloping both C- and O-types of CO\textsubscript{2}RR intermediates
    was introduced to construct volcano plots and limiting potential periodic table,
    aiming for intuitive screening of catalyst candidates for efficient CO\textsubscript{2} reduction to CH\textsubscript{4}.
    The eDOS occlusion experiments were performed to unravel individual orbital contribution to adsorption energy.
    To explore the electronic scale principle governing practical engineering catalytic CO\textsubscript{2}RR activity,
    orbitalwise eDOS shifting experiments based on CNN model were employed.
    The study involves examining the adsorption energy and,
    consequently, catalytic activities while varying supported single atoms.
    This work offers a tangible framework to inform both theoretical screening and experimental synthesis,
    thereby paving the way for systematically designing efficient SACs.
\end{abstract}

\newpage
\section{Introduction}

    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig1_pipeline.png}
        \caption{\textbf{Catalyst performance analysis pipeline integrating volcano plots and CNN.}
        The figure demonstrates the integrated usage of volcano plots for
        predictive assessment of existing catalysts, and the utility of CNNs
        to predict and modulate adsorption energies \(E_{\text{ads}}\) using eDOS as input.
        eDOS occlusion and shifting experiments are conducted to extract chemical information from CNN model.}
        \label{main_fig1:pipeline}
    \end{figure}

@@

\subsection{Adsorption energy prediction from electronic density of states with convolutional neural network}

    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig2_cnn.png}
        \caption{\textbf{CNN for electronic density of states.}
        \textbf{a}. Kendall rank correlation coefficient map illustrating the correlation
        between adsorption energy and other electronic and elementary descriptors.
        Definitions of the notations are provided in
        Supplementary \cref{supp_table15:descriptor_notations}.
        \textbf{b}-\textbf{c}. Architectural of (\textbf{b}) the root and
        (\textbf{c}) each branch of the CNN model.
        The architecture stays consistent across individual branches.
        \textbf{d}-\textbf{m}. Parity plots demonstrating the performance of the
        CNN model across nine adsorbates:
        (\textbf{d}) all adsorbates, (\textbf{e}) CO\textsubscript{2}, (\textbf{f}) COOH,
        (\textbf{g}) CO, (\textbf{h}) CHO, (\textbf{i}) CH\textsubscript{2}O,
        (\textbf{j}) OCH\textsubscript{3}, (\textbf{k}) O, (\textbf{l}) OH, and (\textbf{m}) H.
        The augmentation data are not displayed in these plots, nor included in MAE evaluations.
        Kernel density estimate results are shown along the top and right sides of the main plot,
        representing DFT calculations and CNN model prediction distributions, respectively.}
        \label{main_fig2:cnn_for_eads}
    \end{figure}

@@

\subsection{Hybrid descriptors enabled volcano plot for catalyst analysis}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{main_fig3_volcano.png}
    \caption{\textbf{Limiting potential volcano plot and periodic table.}
    \textbf{a}. Limiting potential volcano plot computed from hybrid-descriptor-based
    linear scaling relations at potential of -0.17 V vs SHE.
    In the plot, the circle, triangle, square, diamond, plus, and star symbols represent
    metals supported on g-C\textsubscript{3}N\textsubscript{4}, nitrogen-doped graphene,
    graphene with dual-vacancy, black phosphorous, BN, and C\textsubscript{2}N, respectively.
    Dotted purple lines delineate regions representing rate-determining steps,
    with roman numerals denoting the respective RDS for each region.
    The reaction steps are defined in \cref{main_eq1:co2rr1} to \cref{main_eq8:co2rr8}.
    \textbf{b}. Limiting potential periodic table illustrating the theoretical activity of
    single metal atom catalysts supported on g-C\textsubscript{3}N\textsubscript{4}.}
    \label{main_fig3:volcano}
\end{figure}

@@

\subsection{Validation of CNN predictions}

    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig4_validation.png}
        \caption{\textbf{Validation of CNN predictions.}
        \textbf{a}-\textbf{c}, The eDOS occlusion experiment results for
        (\textbf{a}) Cr of Cr@g-C\textsubscript{3}N\textsubscript{4},
        (\textbf{b}) Co of Co@g-C\textsubscript{3}N\textsubscript{4} and
        (\textbf{c}) Cu of Cu@g-C\textsubscript{3}N\textsubscript{4}.
        eDOS occlusion experiments were conducted on the initial states of CO adsorption process.
        \textbf{d}-\textbf{f}, Real-space wavefunction visualization for
        (\textbf{d}) Cr $d_{x^2-y^2}$ orbital of Cr@g-C\textsubscript{3}N\textsubscript{4},
        (\textbf{e}) Co $d_{xy}$ orbital of Co@g-C\textsubscript{3}N\textsubscript{4} and
        (\textbf{f}) Cu $d_{xy}$ orbital of Cu@g-C\textsubscript{3}N\textsubscript{4}.
        Wavefunction visualizations were performed on the final states of CO adsorption process.
        The yellow and aqua colors represent areas of positive and negative
        Kohn-Sham orbital coefficients, respectively.
        \textbf{g}-\textbf{j}, Occlusion experiment results for (\textbf{g}) $d_{xy}$ orbital
        and (\textbf{h}) $d_{xz}$ orbital for Co of Co@g-C\textsubscript{3}N\textsubscript{4}.
        COHP analysis for (\textbf{i}) $d_{xy}$ orbital and (\textbf{j}) $d_{xz}$ orbital for
        Co of Co@g-C\textsubscript{3}N\textsubscript{4}, performed between Co and
        all adsorbate atoms in the final states of CO adsorption process.}
        \label{main_fig4:validation}
    \end{figure}

@@

    \subsection{Shifting experiments for prediction of better catalysts}

    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig5_shifting.png}
        \caption{\textbf{eDOS shifting experiments.}
        Impact of orbitalwise eDOS shifting on CO adsorption energy, as predicted by the CNN model,
        for single metal atom catalysts supported on g-C\textsubscript{3}N\textsubscript{4}.
        The disturbances caused by shifting of (\textbf{a}) entire $d$, (\textbf{b}) $d_{xy}$,
        (\textbf{c}) $d_{yz}$, (\textbf{d}) $d_{z^2}$, (\textbf{e}) $d_{xz}$ and
        (\textbf{f}) $d_{x^2-y^2}$ orbitals are illustrated.
        The shifting step size corresponds to the energy resolution of the eDOS,
        and is 0.005 eV in this study.
        A positive eDOS shift indicates a shift towards higher energy levels, and vice versa.}
        \label{main_fig5:shifting}
    \end{figure}

@@

\newpage
\section{Methods}

@@

    \subsection{CO\textsubscript{2}RR pathway}
    The C1 production CO\textsubscript{2}RR pathway, as initially described by Peterson et al. \cite{peterson2010copper} and later confirmed by Jiao et al. \cite{jiao2017molecular} for g-C\textsubscript{3}N\textsubscript{4} based SACs, involves the following elementary steps.
    Asterisks (*) indicate species adsorbed on the SAC surfaces:

@@

\appendix
\newpage
    \section{Code and Data Availability}
    Essential source code and data required to replicate the study's results-including VASP input files,
    structure files of SACs, eDOS arrays and the pretrained CNN model-are available on GitHub repository
    at \url{https://github.com/DanielYang59/cnn4dos}, and can be provided by the corresponding authors upon request.
    Due to their substantial size in terabytes, the complete source data (complete VASP output files and such)
    is not currently hosted on publicly available repositories,
    but can be obtained from the corresponding authors upon request.

@@

\typeout{get arXiv to do 4 passes: Label(s) may have changed. Rerun}

>>>>

% Main
\documentclass[a4paper, 12pt, titlepage]{article}

\usepackage{amsmath}
\usepackage{array}
\usepackage{bookmark}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{cleveref}
\usepackage{datetime2}
\usepackage{enumitem}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{nicematrix}
\usepackage{tablefootnote}
\usepackage{textcomp}
\usepackage{siunitx}
\usepackage{xr}
\usepackage[affil-it]{authblk}
\usepackage[english]{babel}
\usepackage[version=4]{mhchem}

@@

\externaldocument{supplement}

@@

% Title, Authors and Affiliations
\newpage
    \title{Convolutional Neural Networks and Volcano Plots:
    Screening and Prediction of Two-Dimensional Single-Atom Catalysts
    for CO\textsubscript{2} Reduction Reactions}

@@

% Abstract
\newpage
\begin{abstract}
    Single-atom catalysts (SACs) have emerged as pivotal frontiers for catalyzing a myriad of chemical reactions,
    yet the diverse combinations of active elements and support materials, the nature of coordination environments,
    elude traditional methodologies in searching optimal SAC systems with superior catalytic performance.
    Herein, by integrating multi-branch Convolutional Neural Network (CNN) analysis models
    to hybrid descriptor based activity volcano plot,
    two-dimensional (2D) SAC system composed of diverse metallic single atoms anchored on six type of 2D supports,
    including graphitic carbon nitride (g-C\textsubscript{3}N\textsubscript{4}), nitrogen-doped graphene,
    graphene with dual-vacancy, black phosphorous, boron nitride (BN), and C\textsubscript{2}N,
    are screened for efficient CO\textsubscript{2}RR.
    Starting from establishing a correlation map between the adsorption energies of
    intermediates and diverse electronic and elementary descriptors,
    sole singular descriptor lost magic to predict catalytic activity, including d-band center.
    Deep learning method utilizing multi-branch CNN model therefore was employed,
    using 2D electronic density of states (eDOS) as input to predict adsorption energies.
    Hybrid-descriptor enveloping both C- and O-types of CO\textsubscript{2}RR intermediates
    was introduced to construct volcano plots and limiting potential periodic table,
    aiming for intuitive screening of catalyst candidates for efficient CO\textsubscript{2} reduction to CH\textsubscript{4}.
    The eDOS occlusion experiments were performed to unravel individual orbital contribution to adsorption energy.
    To explore the electronic scale principle governing practical engineering catalytic CO\textsubscript{2}RR activity,
    orbitalwise eDOS shifting experiments based on CNN model were employed.
    The study involves examining the adsorption energy and,
    consequently, catalytic activities while varying supported single atoms.
    This work offers a tangible framework to inform both theoretical screening and experimental synthesis,
    thereby paving the way for systematically designing efficient SACs.
\end{abstract}

% Introduction
\newpage
\section{Introduction}

    % Figure 1: Catalyst Performance Analysis Pipeline
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig1_pipeline.png}
        \caption{\textbf{Catalyst performance analysis pipeline integrating volcano plots and CNN.}
        The figure demonstrates the integrated usage of volcano plots for
        predictive assessment of existing catalysts, and the utility of CNNs
        to predict and modulate adsorption energies \(E_{\text{ads}}\) using eDOS as input.
        eDOS occlusion and shifting experiments are conducted to extract chemical information from CNN model.}
        \label{main_fig1:pipeline}
    \end{figure}

@@

% Results Part One: Adsorption energy prediction from eDOS with CNN
\subsection{Adsorption energy prediction from electronic density of states with convolutional neural network}

    % Figure 2: CNN for eDOS
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig2_cnn.png}
        \caption{\textbf{CNN for electronic density of states.}
        \textbf{a}. Kendall rank correlation coefficient map illustrating the correlation
        between adsorption energy and other electronic and elementary descriptors.
        Definitions of the notations are provided in
        Supplementary \cref{supp_table15:descriptor_notations}.
        \textbf{b}-\textbf{c}. Architectural of (\textbf{b}) the root and
        (\textbf{c}) each branch of the CNN model.
        The architecture stays consistent across individual branches.
        \textbf{d}-\textbf{m}. Parity plots demonstrating the performance of the
        CNN model across nine adsorbates:
        (\textbf{d}) all adsorbates, (\textbf{e}) CO\textsubscript{2}, (\textbf{f}) COOH,
        (\textbf{g}) CO, (\textbf{h}) CHO, (\textbf{i}) CH\textsubscript{2}O,
        (\textbf{j}) OCH\textsubscript{3}, (\textbf{k}) O, (\textbf{l}) OH, and (\textbf{m}) H.
        The augmentation data are not displayed in these plots, nor included in MAE evaluations.
        Kernel density estimate results are shown along the top and right sides of the main plot,
        representing DFT calculations and CNN model prediction distributions, respectively.}
        \label{main_fig2:cnn_for_eads}
    \end{figure}

@@

% Results Part Two: Hybrid descriptors enabled volcano plot for catalyst analysis
\subsection{Hybrid descriptors enabled volcano plot for catalyst analysis}

    % Figure 3: Limiting Potential Volcano Plot and Periodic Table
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig3_volcano.png}
        \caption{\textbf{Limiting potential volcano plot and periodic table.}
        \textbf{a}. Limiting potential volcano plot computed from hybrid-descriptor-based
        linear scaling relations at potential of -0.17 V vs SHE.
        In the plot, the circle, triangle, square, diamond, plus, and star symbols represent
        metals supported on g-C\textsubscript{3}N\textsubscript{4}, nitrogen-doped graphene,
        graphene with dual-vacancy, black phosphorous, BN, and C\textsubscript{2}N, respectively.
        Dotted purple lines delineate regions representing rate-determining steps,
        with roman numerals denoting the respective RDS for each region.
        The reaction steps are defined in \cref{main_eq1:co2rr1} to \cref{main_eq8:co2rr8}.
        \textbf{b}. Limiting potential periodic table illustrating the theoretical activity of
        single metal atom catalysts supported on g-C\textsubscript{3}N\textsubscript{4}.}
        \label{main_fig3:volcano}
    \end{figure}

@@

% Results Part Three: Validation of CNN predictions
\subsection{Validation of CNN predictions}

    % Figure 4: Validation of CNN Predictions
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig4_validation.png}
        \caption{\textbf{Validation of CNN predictions.}
        \textbf{a}-\textbf{c}, The eDOS occlusion experiment results for
        (\textbf{a}) Cr of Cr@g-C\textsubscript{3}N\textsubscript{4},
        (\textbf{b}) Co of Co@g-C\textsubscript{3}N\textsubscript{4} and
        (\textbf{c}) Cu of Cu@g-C\textsubscript{3}N\textsubscript{4}.
        eDOS occlusion experiments were conducted on the initial states of CO adsorption process.
        \textbf{d}-\textbf{f}, Real-space wavefunction visualization for
        (\textbf{d}) Cr $d_{x^2-y^2}$ orbital of Cr@g-C\textsubscript{3}N\textsubscript{4},
        (\textbf{e}) Co $d_{xy}$ orbital of Co@g-C\textsubscript{3}N\textsubscript{4} and
        (\textbf{f}) Cu $d_{xy}$ orbital of Cu@g-C\textsubscript{3}N\textsubscript{4}.
        Wavefunction visualizations were performed on the final states of CO adsorption process.
        The yellow and aqua colors represent areas of positive and negative
        Kohn-Sham orbital coefficients, respectively.
        \textbf{g}-\textbf{j}, Occlusion experiment results for (\textbf{g}) $d_{xy}$ orbital
        and (\textbf{h}) $d_{xz}$ orbital for Co of Co@g-C\textsubscript{3}N\textsubscript{4}.
        COHP analysis for (\textbf{i}) $d_{xy}$ orbital and (\textbf{j}) $d_{xz}$ orbital for
        Co of Co@g-C\textsubscript{3}N\textsubscript{4}, performed between Co and
        all adsorbate atoms in the final states of CO adsorption process.}
        \label{main_fig4:validation}
    \end{figure}

@@

% Results Part Four: Shifting experiments for prediction of better catalysts
\subsection{eDOS shifting experiments for prediction of better catalysts}

    % Figure 5: eDOS Shifting Experiments
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig5_shifting.png}
        \caption{\textbf{eDOS shifting experiments.}
        Impact of orbitalwise eDOS shifting on CO adsorption energy, as predicted by the CNN model,
        for single metal atom catalysts supported on g-C\textsubscript{3}N\textsubscript{4}.
        The disturbances caused by shifting of (\textbf{a}) entire $d$, (\textbf{b}) $d_{xy}$,
        (\textbf{c}) $d_{yz}$, (\textbf{d}) $d_{z^2}$, (\textbf{e}) $d_{xz}$ and
        (\textbf{f}) $d_{x^2-y^2}$ orbitals are illustrated.
        The shifting step size corresponds to the energy resolution of the eDOS,
        and is 0.005 eV in this study.
        A positive eDOS shift indicates a shift towards higher energy levels, and vice versa.}
        \label{main_fig5:shifting}
    \end{figure}

@@

% Methods
\newpage
\section{Methods}

@@

    \subsection{CO\texorpdfstring{\textsubscript{2}}{2}RR pathway}
    The C1 production CO\textsubscript{2}RR pathway, as initially described by Peterson et al. \cite{peterson2010copper} and later confirmed by Jiao et al. \cite{jiao2017molecular} for g-C\textsubscript{3}N\textsubscript{4} based SACs, involves the following elementary steps.
    Asterisks (*) indicate species adsorbed on the SAC surfaces:

@@

% Code/Data Availability, Acknowledgements, Author Contributions and Competing Interests
\appendix
\newpage
    \section{Code and Data Availability}
    Essential source code and data required to replicate the study's results-including VASP input files,
    structure files of SACs, eDOS arrays and the pretrained CNN model-are available on GitHub repository
    at \url{https://github.com/DanielYang59/cnn4dos}, and can be provided by the corresponding authors upon request.
    Due to their substantial size in terabytes, the complete source data (complete VASP output files and such)
    is not currently hosted on publicly available repositories,
    but can be obtained from the corresponding authors upon request.

@@
